A linear programming approach to the tracking of partials

### 2017-04-24 paper/paper.tex
Improved plots, wrote about example in paper.

We compare the greedy and LP based methods for peak matching on a synthetic
signal. The signal is composed of $Q=6$ chirps of constant amplitude, the $q$th
chirp $s$ at sample $n$ described by the equation
\[
    s_{q}(n) = \exp(j(\phi_{q} + \omega_{q}n +
    \frac{1}{2} \psi_{q} n^{2}))
\]
The parameters for the 6 chirps are presented in
Table~\ref{tab:ptrackexamplechirpparams}.

@@

Two 1 second long signals are synthesized at a sampling rate of 16000 Hz, the
first with chirps 0--2, the second with chirps 3--5. We add
Gaussian distributed white noise at several SNRs to evaluate the technique in the
presence of noise.

%\begin{figure}[!t]
%    %\centering
%    \centering
%    \includegraphics[width=\figwidthscale\textwidth]{plots/mq_lp_compare_chirp_20.eps}
%    \CaptionWithTitle{%
%        \input{plots/mq_lp_compare_chirp_20.txt}%
%    }{ Line-segments representing the frequency and frequency-slope at local
%        spectrogram maxima. In the bottom two plots the line segments not deemed
%        by the respective algorithms as belonging to a partial path are
%        discarded, revealing the estimated partial trajectories. See
%        Table~\ref{tab:ptrackexamplechirpparams} for the chirp parameters.
%    \label{plot:mq_lp_compare_chirp_20}}
%\end{figure}
%\begin{figure}[!t]
%    %\centering
%    \centering
%    \includegraphics[width=\figwidthscale\textwidth]{plots/mq_lp_compare_chirp_15.eps}
%    \CaptionWithTitle{%
%        \input{plots/mq_lp_compare_chirp_15.txt}%
%    }{ Line-segments representing the frequency and frequency-slope at local
%        spectrogram maxima. In the bottom two plots the line segments not deemed
%        by the respective algorithms as belonging to a partial path are
%        discarded, revealing the estimated partial trajectories. See
%        Table~\ref{tab:ptrackexamplechirpparams} for the chirp parameters.
%    \label{plot:mq_lp_compare_chirp_15}}
%\end{figure}
\begin{figure}[!t]
    %\centering
    \centering
    \centerline{\includegraphics[width=\figwidthscale\textwidth]{plots/mq_lp_compare_chirp_10.eps}}
    \CaptionWithTitle{%
        \input{plots/mq_lp_compare_chirp_10.txt}%
    }{ Line-segments representing the frequency and frequency-slope at local
        spectrogram maxima. In the bottom two plots the line segments not deemed
        by the respective algorithms as belonging to a partial path are
        discarded, revealing the estimated partial trajectories. See
        Table~\ref{tab:ptrackexamplechirpparams} for the chirp parameters.
    \label{plot:mq_lp_compare_chirp_10}}
\end{figure}

A spectrogram of each signal is computed with an analysis window length of 1024
samples and a hop-size $H$ of 256 samples. Local maxima are searched in 150 Hz
wide bands spaced 75 Hz apart. A local maximum is only accepted if its amplitude
is greater than -20 dB. The bin corresponding to each local maximum and its two
surrounding bins are used by the Distribution Derivative Method
(DDM) to estimate the
local chirp parameters, the $i$th set of parameters in frame $k$ denoted
$\theta_{i}^{k} = \left\{ \phi_{i}^{k} , \omega_{i}^{k} , \psi_{i}^{k}
\right\}$ (the atoms used by the DDM are generated from 4-term continuous
Nuttall windows). The results of the analyses of both signals are lumped together and
it is on this lumped data that we perform partial tracking.

We search for partial tracks using both the greedy and LP strategies. Both
algorithms use the distance metric $\mathcal{D}_{\text{pr.}}$ between two parameters sets:
\[
    \mathcal{D}_{\text{pr.}} \left( \theta_{i}^{k},
    \theta_{j}^{k+1} \right) = \left( \omega_{i}^{k} +
    \psi_{i}^{k} H - \omega_{j}^{k+1} \right)
\]
which is the error in predicting $j$th frequency in frame $k+1$ from the $i$th
parameters in frame $k$. For the greedy method, the search for partial paths is
restricted to one frame ahead like in \cite{mcaulay1986speech}, otherwise the
computation becomes intractable. For the LP
method, to keep the computation time reasonable, we search over 6 frames for 6
best paths (the number of paths does not affect the computation time).
To maintain connected paths, the search on the next frames uses the end nodes of
the last search as starting points. For both methods, the search is restricted
to nodes between frequencies 250 to 2250 Hz.

Figure~\ref{plot:mq_lp_compare_chirp_10}
shows discovered partial trajectories for signals with a SNR of 10
dB. It is seen that while the greedy method performs poorly
at a SNR of 10dB, the LP method still gives plausible partial trajectories.

@@

There are problems with the proposed approach. Perhaps ``jagged'' paths should be removed using regularization.
There are also situations where it is undesirable to have paths extend
throughout the entire lattice. Acoustic signals produced by striking media, such
as strings or bars, exhibit a spectrum where the upper partials decay more
quickly than the lower ones --- it
would be desirable in these situations to have shorter paths for the upper
partials, those decaying more quickly. This could be addressed as in
\cite{depalle1993tracking} where the signal is divided into overlapping sequences of
frames and partial paths are connected between sequences.

The proposed algorithm, while faster than algorithms based on the Viterbi
algorithm, is still not fast. In many situations where computational resources are
limited, a McAulay-Quatieri method search over many sets of small $K$ works
sufficiently well. However in high amounts of noise the algorithm proposed here
is robust and of tractable complexity.

>>>>

We compare the greedy and LP based methods for peak matching on a synthetic
signal. The signal is composed of $Q=3$ chirps of constant amplitude, the $q$th
chirp $s$ at sample $n$ described by the equation
\[
    s_{q}(n) = \exp(j(\phi_{q} + \omega_{q}n +
    \frac{1}{2} \psi_{q} n^{2}))
\]
The parameters for the $Q$ chirps are presented in
Table~\ref{tab:ptrackexamplechirpparams}.

@@

A 1 second long signal is synthesized at a sampling rate of 16000 Hz, the
chirps ramping from their initial to final frequency in that time. We add
Gaussian distributed white noise at several SNRs to evaluate the technique in the
presence of noise.

\begin{figure*}
    %\centering
    \centering
    \centerline{\includegraphics[width=0.75\textwidth]{plots/mq_lp_compare_chirps.eps}}
    { Line-segments representing the frequency and frequency-slope at local
        spectrogram maxima. The coloured segments correspond to connected paths
        returned by the search algorithms. Plots 1--3.a. show the atomic
        decomposition of the signal before partial tracking for SNRs of 0, -6
        and -12 dB, respectively. Plots 1--3.b. show the partial paths
        discovered by the LP method and plots 1--3.c. show the paths discovered
        by the McAulay-Quatieri method. 
        See Table~\ref{tab:ptrackexamplechirpparams} for the chirp parameters.
    \label{plot:mq_lp_compare_chirps}}
\end{figure*}

A spectrogram of each signal is computed with an analysis window length of 2048
samples and a hop-size $H$ of 512 samples. Local maxima are searched in 100 Hz
wide bands spaced 50 Hz apart. The bin corresponding to each local maximum and its two
surrounding bins are used by the Distribution Derivative Method
(DDM) \cite{betser2009sinusoidal} to estimate the
local chirp parameters, the $i$th set of parameters in frame $k$ denoted
$\theta_{i}^{k} = \left\{ \phi_{i}^{k} , \omega_{i}^{k} , \psi_{i}^{k}
\right\}$ (the atoms used by the DDM are generated from 4-term continuous
Nuttall windows \cite{nuttall1981some}). Partial tracking is performed on the
resulting atomic decomposition.

We search for partial tracks using both the greedy and LP strategies. Both
algorithms use the distance metric $\mathcal{D}_{\text{pr.}}$ between two parameters sets:
\[
    \mathcal{D}_{\text{pr.}} \left( \theta_{i}^{k},
    \theta_{j}^{k+1} \right) = \left( \omega_{i}^{k} +
    \psi_{i}^{k} H - \omega_{j}^{k+1} \right)
\]
which is the error in predicting $j$th frequency in frame $k+1$ from the $i$th
parameters in frame $k$. For the greedy method, the search for partial paths is
restricted to two frames ahead, i.e., paths of length $K_{\text{MQ}}=3$ are sought, otherwise the
computation becomes intractable. For the LP
method the search is carried out over all frames ($K_{\text{LP}}=28$). %TODO: Is it?
The cost thresholding values are ${\Delta_{\text{MQ}} = \Delta_{\text{LP}} =
0.1}$. For both methods, the search is restricted
to nodes between frequencies 250 to 2000 Hz.

Figure~\ref{plot:mq_lp_compare_chirps}
shows discovered partial trajectories for signals at various SNRs. It is seen
that while the greedy method starts performing poorly at an SNR of -6db, the LP method
still gives plausible partial trajectories. The LP method returns paths spanning
all $K$ frames, due to the constraints. The McAulay-Quatieri method in general does not,
but longer paths can be formed in a straightforward way after the initial short path
search step \cite{mcaulay1986speech}.

It is interesting to note that the paths are found by only considering the
prediction error of the initial frequency of the atom. Other cost functions can
be chosen depending on the nature of the signal: reasonable cost functions here
might be similarity of the atoms's energies or frequency slopes.

@@

There are problems with the proposed approach. 
There are situations where it is undesirable to have paths extend
throughout the entire lattice. Acoustic signals produced by striking media, such
as strings or bars, exhibit a spectrum where the upper partials decay more
quickly than the lower ones; it
would be desirable in these situations to have shorter paths for these
partials. This could be addressed as in
\cite{depalle1993tracking} where the signal is divided into overlapping sequences of
frames and partial paths are connected between sequences.

The proposed algorithm, while faster than algorithms based on the Viterbi
algorithm \cite{wolf1989finding}, is still not fast. In many situations where computational resources are
limited, a McAulay-Quatieri method search over many sets of small $K$ works
sufficiently well. However in high amounts of noise the algorithm proposed here
is robust while still of tractable complexity.
